Figures to insert in text

### 2018-03-08 Chapters/Chapter02.tex
sdm conclusion

@@

Performance is an important part of our framework. Running an SDM memory consumes a lot of resource from both CPU and RAM.

For instance, each scan in 1,000,000 hard-locations with 1,000 bits executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount. So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares. The heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} executed $3.05 \cdot 10^{15}$ bit compares. For comparison, the number of seconds since Jesus's birth is 63,639,648,000 = $6.36 \cdot 10^{10}$. The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$. There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.

Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.

Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.

Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds. It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.

We have executed the same performance test in each device. The test has 3 parts. The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans. The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners. The code is available in the ``Performance test'' notebook \citep{sdmframework}.

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
\label{fig:performance-imac-scan-1000}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
\label{fig:performance-imac-scan-256}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
\label{fig:performance-imac-scan-10k}}
\end{figure}

\begin{center}
    iMac $n=1,000$ bits
    \begin{tabular}{| l | r | r | r |}
    \hline
    & Loops & Scans / second & Scan time (ms) \\ \hline
    Linear scan & 1000 & 81.62 & 12.25 \\
    Thread scan & 5000 & 143.68 & 6.95 \\
    OpenCL scan & 5000 & 238.00 & 4.20 \\ \hline
    \hline
    & Loops & Ops / second & Op. time (ms) \\ \hline
    Thread write & 1000 & 74.92 & 13.34 \\
    Thread single read & 1000 & 96.22 & 10.39 \\
    OpenCL write & 1000 & 126.50 & 7.90 \\
    OpenCL single read & 1000 & 190.20 & 5.25 \\
    \hline
    \end{tabular}
\end{center}

\begin{center}
    iMac $n=256$ bits
    \begin{tabular}{| l | r | r | r |}
    \hline
    & Loops & Scans / second & Scan time (ms) \\ \hline
    Linear scan & 1000 & 175.48 & 5.69 \\
    Thread scan & 5000 & 352.63 & 2.83 \\
    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
    \hline
    & Loops & Ops / second & Op. time (ms) \\ \hline
    Thread write & 2000 & 304.46 & 3.28 \\
    Thread single read & 2000 & 391.21 & 2.55 \\
    OpenCL write & 2000 & 378.44 & 2.64 \\
    OpenCL single read & 2000 & 466.16 & 2.14 \\
    \hline
    \end{tabular}
\end{center}

\begin{center}
    iMac $n=10,000$ bits
    \begin{tabular}{| l | r | r | r |}
    \hline
    & Loops & Scans / second & Scan time (ms) \\ \hline
        Linear scan & 100 & 8.59 & 116.38 \\
        Thread scan & 500 & 18.66 & 53.56 \\
        OpenCL scan & 1000 & 77.20 & 12.95 \\
    \hline
    \end{tabular}
\end{center}

@@

@@

A reduction from measuring the distance to $2^{20}$ hard locations to a computation of $2^{10}$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.  At large scales of computing, the very ability to perform some experiments is a function of sheer performance. The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}. More on this in my closing words.

@@

It is no coincidence that a scientific journal such as Neurocomputing states that “software is scientific method by machine”.  Indeed, recently, Nature analysed the top-100 cited papers in history, to find

\begin{quote}
... some surprises, not least that it takes a staggering 12,119 citations to rank in the top 100 — and that many of the world’s most famous papers do not make the cut. A few that do, such as the first observation of carbon nanotubes (number 36) are indeed classic discoveries. But the vast majority describe experimental methods or software that have become essential in their fields. [...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets. \\
\hfill --- \citet{van2014top}
\end{quote}

The overarching intention here is to not only provide a starting point, but to provide the Framework in which SDM research can be conducted.  Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous best, under a particular benchmark set.  For example, some of the benchmarks that we plan to develop for future research is: how fast is convergence through iterative reading?  How large is the attractor of the critical distance?  How well does the system filter noise?  How well does the system work under the aforementioned supervised learning task?  And other authors may be able to improve this benchmark set themselves, as is usual in open source development.  It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.

Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).  We have referred to this modified read operation as Chada read.  Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.

Of course, there is also new work here.  The mathematics of the model has been shown to be correct (with a single exception); we have shown how to execute unsupervised learning with nothing besides operations originally proposed by Kanerva; we have studied the generalized Murilo read; we have seen and finally, we have reproduced

\subsection{From theory to a platform.}

Platforms in the history of computing.

Opens 10 doors.

@@

\chapter*{List of jupyter notebooks}

>>>>

% ?????????  In Figure

@@

Our intention is to provide comparative performance metrics under different computation engines (CPU, GPU, etc) and different operating systems (Linux, MacOs, Windows, etc). Performance can be measured as the average number of scans of all hard locations per second, reads per second, writes per second, etc.

Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.  We also intend to test on machines such as the iMac with dedicated GPU, MacPro with dedicated GPU, and personal computers under Linux with dedicated GPUs.

Beyond that, we are running as state-of-the-art devices: (i) an Amazon EC2 p3.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.8xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and 8x NVIDIA K80 GPU.

@@

Deep neural networks, alphaGo, alphazero

@@

A reduction from measuring the distance to $2^{20}$ hard locations to a computation of $2^{10}$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.  At large scales of computing, the very ability to perform some experiments is a function of sheer performance. The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}. A little more on this in my closing words.

@@

The overarching intention here is to not only provide a starting point, but to provide a Framework in which SDM research can be conducted.  Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).  For example, some of the benchmarks that we plan to develop in future research is: how fast is convergence through iterative reading?  How large is the attractor of the critical distance?  How well does the system filter noise?  How well does the system work under the supervised learning task?  And other authors may be able to improve this benchmark set themselves, as is usual in open source development.  It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.

Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).  We have referred to this modified read operation as Chada read (the tale is that my friend \text{\&} colleague, Dr. Daniel de Magalhães Chada, along with Linhares, did not consult Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterwards).  Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.  Indeed, a high-quality codebase seems to have become a scientific community's form of unequivocally standing behind a consensus. For example, the journal Nature analysed the top-100 cited papers in history, to find:

\begin{quote}
... some surprises, not least that it takes a staggering 12,119 citations to rank in the top 100 — and that many of the world’s most famous papers do not make the cut. A few that do, such as the first observation of carbon nanotubes (number 36) are indeed classic discoveries. But the vast majority describe experimental methods or \emph{software that have become essential in their fields. [...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.} \\
\hfill --- \citet{van2014top}, emphasis mine.
\end{quote}

It is no coincidence that scientific journals such as BMC Neuroscience, or the Journal of Machine Learning Research have specific sections on open-source software. Neurocomputing states that “software is scientific method by machine”.

Of course, for the skeptical reader who may consider software a less worthy pursuit, there is also new work here.  The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.  The emphasis might have been on \emph{breadth of topics}, in detriment of depth here or there.  But this is due to our research group's enthusiam for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.  If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.  The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate. Some essays completely shut the door close at the end; this one intends to leave it wide open. As the reader may have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.

\begin{figure}[!p]
\centering

\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}

\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}

\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}

\caption{Some slides from an old Linhares class \citet{only-one-life}.  In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.  It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.  Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
\label{fig:one-life}}
\end{figure}

@@
